---
title: "Comparison of ICES data-rich and data-limited methods with MSE"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

This repository contains the data and code for a comparison of the data-rich ICES MSY rule to the ICES data-limited empirical methods (rfb rule, hr rule).

Three case study stocks are included:

1.  Plaice (*Pleuronectes platessa*) in Division 7.e (western English Channel)

2.  Cod (Gadus morhua) in Subarea 4, Division 7.d, and Subdivision 20 (North Sea, eastern English Channel, Skagerrak)

3.  Herring (Clupea harengus) in Subarea 4 and divisions 3.a and 7.d, autumn spawners (North Sea, Skagerrak and Kattegat, eastern English Channel)

The operating models (OMs) are created using the SAM [`stockassessment`](https://github.com/fishfollower/SAM/) R package and follows the principles developed during the ICES Workshop on North Sea stocks management strategy evaluation ([WKNSMSE](https://doi.org/10.17895/ices.pub.5090)).

The simulation is based on the Fisheries Library in R ([FLR](http://www.flr-project.org/)) and its `mse` package.

The data-limited MSE framework is an adaptation of Fischer et al. ([2021](https://dx.doi.org/10.1093/icesjms/fsab169)), see the [`shfischer/GA_MSE_PA`](https://github.com/shfischer/GA_MSE_PA) GitHub repository for the original source code and documentation.

For exact reproducibility, R packages versions are recorded with [renv](https://rstudio.github.io/renv/) in a [renv.lock](https://github.com/shfischer/MSE_risk_comparison/blob/master/renv.lock) file.

## Repository structure

-   `funs_*`: Function libraries, defining the functions used in the other scripts

    -   `funs.R`: generic function library, including definition of data-limited management procedures (MPs)

    -   `funs_analysis.R`: for analysis of results

    -   `funs_GA.R`: functions used in the optimisation with the genetic algorithm (GA)

    -   `funs_OM.R`: functions for generating the operating models

    -   `funs_WKNSMSE.R`: functions required for the ICES MSY rule

-   `OM_*`: Scripts for operating models (OMs, including alternative OMs)

    -   `OM_ple.27.7e.R` for plaice

    -   `OM_cod.27.47d20.R` for cod

    -   `OM_her.27.3a47d.R` for herring

    -   `OM_MSY.R`: script for estimating MSY

    -   `OM_MSY.pbs`: job script, calling `OM_MSY.R`

-   `MP_*`: Script for running and analysing the MSE

    -   `MP_analysis.R`: script for analysing MSE results (summarising, exporting, visualisation, ...)

    -   `MP_run.R`: script for running any MP in the MSE and optimising MPs

    -   `MP_*.pbs`: job submission scripts, used for running MP_run.R on a high-performance computing cluster, e.g. `MP_run_rfb_mult.pbs` for optimising the rfb rule with a multiplier

-   `OM*.R`: Scripts for creating the operating model (OM)

    -   `OM_SAM_fit.R` script for configuring and running SAM,
    -   `OM_Eqsim.R` runs EqSim on SAM model fit to create ICES style reference points,
    -   `OM_length.R` preparation of length data for the generation of the length index for the rfb rule,
    -   `OM.R` generate the operating model(s),
    -   `OM_MSY.R` calculation of real MSY reference points with the mse framework

`input/`: This directory contains the files required for generating the OMs for the three stocks

-   `input/ple.27.7e/preparation/`: for plaice

-   `input/cod.27.47d20/preparation/`: for cod

-   `input/her.27.3a47d/preparation/`: for herring

-   `input/OM_refpts.csv`: summarised OM reference points

`output/`: This directory contains some summarised results

-   `output/MPs_baseline.csv`: summary statistics and parameterisations of default and optimised MP for all stocks

-   [`output/plots/wormplots/`](https://github.com/shfischer/MSE_risk_comparison/tree/master/output/plots/wormplots): Projections (wormplots) for all default and optimised MPs, for all stocks and OMs

## R, R packages and version info

The MSE simulations were run with R:

```{r, eval = FALSE}
> sessionInfo()
R version 4.1.0 (2021-05-18)
Platform: x86_64-w64-mingw32/x64 (64-bit)
Running under: Windows 10 x64 (build 19041)
```

The package versions and their dependencies are recorded with the R package [renv](https://rstudio.github.io/renv/) and stored in the file [renv.lock](https://github.com/shfischer/MSE_risk_comparison/blob/master/renv.lock). The exact package version can be restored by cloning this repository, navigating into this folder in R (or setting up a project), installing the renv package

```{r, eval = FALSE}
install.packages("renv")
```

and calling

```{r, eval = FALSE}
renv::restore()
```

See [renv](https://rstudio.github.io/renv/) and the package documentation for details.

The framework is based on the Fisheries Library in R (FLR) framework and uses the [FLR packages](https://flr-project.org/) `FLCore`, `FLasher`, `FLBRP`, `FLAssess`, `FLXSA`, `ggplotFL`, `mse`, and `FLfse`. See [renv.lock](https://github.com/shfischer/MSE_risk_comparison/blob/master/renv.lock) for version details and sources.

Also, the R package [`stockassessment`](https://github.com/fishfollower/SAM)is used.

For running the optimisations on a high-performance computing cluster, a suitable MPI back-end and the R package `Rmpi` are required.
